---
title: "Innocent Death's caused by Police Project"
subtitle: "Presentation subtitle (if any)"
author: "I Have Never Coded Before<br> Abhey Reuben Mieke"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(here)
library(usdata)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
# Load your data here
police_data <- read_csv(here("data/Innocent Deaths caused by Police (All time).csv"))
head(police_data)

glimpse(head(police_data))

police_data <- rename(police_data, unique_id = `Unique ID`, name = Name, age = Age, gender = Gender, race = Race, doi = `Date of injury resulting in death (month/day/year)`, loi_address = `Location of injury (address)`, state = State, lod_zip_code = `Location of death (zip code)` , lod_county = `Location of death (county)`, latitude = Latitude, longitude = Longitude,  agency_involved = `Agency or agencies involved`, hlof = `Highest level of force`, alleged_weapon = `Alleged weapon`, agressive_movement = `Aggressive physical movement`, fleeing_not_fleeing = `Fleeing/Not fleeing`, description_temp = `Description Temp`, dispositions_exclusions =`Dispositions/Exclusions INTERNAL USE, NOT FOR ANALYSIS`, iuof = `Intended use of force (Developing)`, mental_illness = `Foreknowledge of mental illness`) %>% mutate(state = case_when(state == c(state) ~ abbr2state(c(state)))) %>% mutate(year = str_sub(doi,-4,-1))

#doi = date of injury resulting in death, loi= location of injury resulting in death, lod = location of death, hlof = highest level of force, iuof = intended use of force

police_data <- subset(police_data, select = -c(6, 8, 13, 21, 22, 23, 26))
police_data1 <- police_data %>%
  filter(age < 18)

```

```{r no, include=FALSE}
style_xaringan(
  title_slide_background_image = "img/confetti.jpg"
)
```

class: center, middle

To see if there is a correlation between dying by the police and being innocent. We started off the presentation thinking there's no correlation.

---

class: inverse, center, middle

# Section title

---

# Hello World

- Click the `Knit` button to compile your presentation

- Make sure to commit and push all resulting files to your GitHub repo

---

class: inverse, middle, center

# Using xaringan

---

# xaringan

- The presentation is created using the `xaringan` package

- Use `---` to separate slides and `--` for incremental builds

--

- Like this

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
#iris_modified <- iris %>%
police_time_race <-police_data %>% mutate(year = str_sub(doi,-4,-1)) %>% mutate(race = case_when(
      name == "Oliver Sevin Frazier-Savoy" ~ "European-American/White",
      TRUE ~ race
    )
  ) %>% group_by(year,race) %>% count(year) 

ggplot() + geom_col(police_time_race, mapping = aes(x = year, y = n, fill = race)) + facet_wrap(~ race, nrow = 4) + labs( x = "Year", y = "Number of Deaths", title = "Police Deaths by Race") + scale_x_discrete(guide = guide_axis(angle = 45))

#https://www.census.gov/prod/2001pubs/cenbr01-1.pdf


```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
#ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
 # geom_point() + 
  #theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html

#k + geom_map(aes(map_id = state), map = map) + expand_limits(x = map$long, y = map$lat)
#map_id, alpha, color, fill, linetype, size

police_data_hm <- police_data %>% 
  filter(year == 2019) %>% 
  count(state) %>% 
  mutate(n = case_when(
    state == "California" ~ 212/(330150668*0.1196),
    state == "Texas" ~ 178/(330150668*0.0868),
    state == "Florida" ~ 129/(330150668*0.0644),
    state == "New York" ~ c(n)/(330150668*0.0591),
    state == "Pennsylvania" ~ c(n)/(330150668*0.0387),
    state == "Illinois" ~ c(n)/(330150668*0.0385),
    state == "Ohio" ~ c(n)/(330150668*0.0353),
    state == "Georgia" ~ c(n)/(330150668*0.0318),
    state == "North Carolina" ~ c(n)/(330150668*0.0314),
    state == "Michigan" ~ c(n)/(330150668*0.0302),
    state == "New Jersey" ~ c(n)/(330150668*0.0269),
    state == "Virginia" ~ c(n)/(330150668*0.0258),
    state == "Washington" ~ c(n)/(330150668*0.0228),
    state == "Arizona" ~ c(n)/(330150668*0.0217),
    state == "Massachusetts" ~ c(n)/(330150668*0.0209),
    state == "Tennessee" ~ c(n)/(330150668*0.0205),
    state == "Indiana" ~ c(n)/(330150668*0.0202),
    state == "Missouri" ~ c(n)/(330150668*0.0185),
    state == "Maryland" ~ c(n)/(330150668*0.0183),
    state == "Wisconsin" ~ c(n)/(330150668*0.0176),
    state == "Colorado" ~ c(n)/(330150668*0.0172),
    state == "Minnesota" ~ c(n)/(330150668*0.0170),
    state == "South Carolina" ~ c(n)/(330150668*0.0154),
    state == "Alabama" ~ c(n)/(330150668*0.0148),
    state == "Louisiana" ~ c(n)/(330150668*0.0141),
    state == "Kentucky" ~ c(n)/(330150668*0.0135),
    state == "Oregon" ~ c(n)/(330150668*0.0127),
    state == "Oklahoma" ~ c(n)/(330150668*0.0119),
    state == "Connecticut" ~ c(n)/(330150668*0.0108),
    state == "Utah" ~ c(n)/(330150668*0.0096),
    state == "Iowa" ~ c(n)/(330150668*0.0095),
    state == "Nevada" ~ c(n)/(330150668*0.0092),
    state == "Arkansas" ~ c(n)/(330150668*0.0091),
    state == "Mississippi" ~ c(n)/(330150668*0.0090),
    state == "Kansas" ~ c(n)/(330150668*0.0088),
    state == "New Mexico" ~ c(n)/(330150668*0.0063),
    state == "Nebraska" ~ c(n)/(330150668*0.0058),
    state == "Idaho" ~ c(n)/(330150668*0.0053),
    state == "West Virginia" ~ c(n)/(330150668*0.0055),
    state == "Hawaii" ~ c(n)/(330150668*0.0043),
    state == "New Hampshire" ~ c(n)/(330150668*0.0041),
    state == "Maine" ~ c(n)/(330150668*0.0040),
    state == "Montana" ~ c(n)/(330150668*0.0032),
    state == "Rhode Island" ~ c(n)/(330150668*0.0032),
    state == "Delaware" ~ c(n)/(330150668*0.0029),
    state == "South Dakota" ~ c(n)/(330150668*0.0027),
    state == "North Dakota" ~ c(n)/(330150668*0.0023),
    state == "Alaska" ~ c(n)/(330150668*0.0022),
    state == "Vermont" ~ c(n)/(330150668*0.0019),
    state == "Wyoming" ~ c(n)/(330150668*0.0017))) %>%
  arrange(desc(n))

#divided number of deaths in a state by percentage of US population living in that state in 2019. Used numbers from https://simple.wikipedia.org/wiki/List_of_U.S._states_by_population
    
states <- map_data("state")
police <- police_data_hm
names(police) <- tolower(names(police))
police$state <- tolower(police_data_hm$state)

choro <- merge(states, police, sort = FALSE, by.x = 'region', by.y = "state")
choro <- choro[order(choro$order), ]
ggplot(choro, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = n)) + 
  labs(fill = "Prop. of Deaths by State Pop.") + 
  theme_minimal() 

  coord_map("albers",  lat0 = 45.5, lat1 = 29.5) 
  

#ggplot(choro, aes(long, lat)) +
#geom_polygon(aes(group = group, fill = assault / murder)) +
#coord_map("albers",  lat0 = 45.5, lat1 = 29.5)
#Code from https://ggplot2.tidyverse.org/reference/map_data.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
police_time_race_proportion <- police_data %>% mutate(year = str_sub(doi,-4,-1)) %>% 
  mutate(
    race = case_when(
      name == "Oliver Sevin Frazier-Savoy" ~ "European-American/White",
      TRUE ~ race
    )
  ) %>% filter(year == 2000) 


```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
police_time_race_proportion <- police_data %>% 
  mutate(year = str_sub(doi,-4,-1)) %>% 
  mutate(
    race = case_when(
      name == "Oliver Sevin Frazier-Savoy" ~ "European-American/White",
      TRUE ~ race
    )
  ) %>% 
  filter(year == 2000 | year == 2010 | year == 2019) %>% filter(race == "African-American/Black" | race == "Asian/Pacific Islander" | race == "European-American/White" | race == 	
"Hispanic/Latino"  | race == 	
"Native American/Alaskan") %>% 
  group_by(year,race)  %>% 
  count(year) %>% 
  mutate(n = case_when(
      year == 2000 & n == 125 ~ 125/34658190,
      year == 2000 & n == 12 ~ 12/10641833,
      year == 2000 & n == 123 ~ 123/211460626,
      year == 2000 & n == 57 ~ 57/35305818,
      year == 2000 & n == 13 ~ 13/2475956,
      year == 2010 & n == 295 ~ 295/38929319,
      year == 2010 & n == 18 ~ 22/15214265,
      year == 2010 & n == 449 ~ 449/223553265,
      year == 2010 & n == 156 ~ 156/ 50477594,
      year == 2010 & n == 13 ~ 13/ 2932248,
      year == 2019 & n == 450 ~ 439/43984096,
      year == 2019 & n == 40 ~ 40/20022610,
      year == 2019 & n == 669 ~ 669/197271953,
      year == 2019 & n == 249 ~ 249/60724311,
      year == 2019 & n == 17 ~ 17/4267113))

ggplot() + geom_col(police_time_race_proportion, mapping = aes(x = race, y = n, fill = race)) + facet_wrap(~ year, nrow = 4) + labs( x = "Race", y = "Proportion of Populations who have Died by Police", title = "Proportion of Populations dying at the hands of the Police") + scale_x_discrete(guide = guide_axis(angle = 45)) 

#
  #data from cencus and middle eastern included within asian race as census does not specify middle eastern race https://www.census.gov/quickfacts/fact/table/US/PST045219, https://www.census.gov/prod/cen2010/briefs/c2010br-02.pdf, https://www.census.gov/quickfacts/fact/table/US/PST045219 calculations rounded down for 2019 as you cannot get a fraction of a person. Also combined population of asian and Native Hawaiian and Other Pacific Islander from cencus. Divided number of deaths by race for each year by population from cencus from that year.
```

```{r innocent_state}
police_data_innocent <- police_data %>%
  filter(alleged_weapon == "None" | is.na(alleged_weapon))  %>%
  filter(agressive_movement == "None" | is.na(agressive_movement)) %>%
  filter(fleeing_not_fleeing == "Not fleeing" | is.na(fleeing_not_fleeing))  %>%
  mutate("firstzip" = str_sub(lod_zip_code, -5,-5)) %>% 
   mutate(year = str_sub(doi,-4,-1)) %>%
  group_by(firstzip, year)





```

---

# Images

```{r wow, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.

---

background-image: url(https://www.google.com/search?q=first+digit+zip+code+map&client=avast-a-1&sxsrf=AOaemvIfjvSPwTFnTZg8yUpp9bTWMuvZCg:1637760678673&source=lnms&tbm=isch&sa=X&ved=2ahUKEwjzs-r7jbH0AhVQ4aQKHbjUB3UQ_AUoAXoECAEQAw&biw=1280&bih=577&dpr=1.5#imgrc=mya_FV0kn9TnwM)

---


```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://www.google.com/search?q=first+digit+zip+code+map&client=avast-a-1&sxsrf=AOaemvIfjvSPwTFnTZg8yUpp9bTWMuvZCg:1637760678673&source=lnms&tbm=isch&sa=X&ved=2ahUKEwjzs-r7jbH0AhVQ4aQKHbjUB3UQ_AUoAXoECAEQAw&biw=1280&bih=577&dpr=1.5#imgrc=mya_FV0kn9TnwM")
```